---
title: "Municipal Data (2022)"
author: "Pietro Colombo"
date: "`r 13/03/2025`"
format: html
---

# **Municipal Data Analysis (2022)**

## **1. General Municipality Data**

The dataset provides details about different municipalities, including population, land, and water areas.

### **Columns:**

-   **COD RIP, COD REG, COD PROV, COD COM**: Coded identifiers for territorial divisions.
-   **COMUNE**: Municipality name.
-   **POP 31/12/2022**: Population count as of December 31, 2022.
-   **SUP 31/12/2022 (Ha)**: Total land area in hectares.
-   **SUP ACQUE (Ha)**: Water area in hectares.

### **Example Data:**

| Municipality    | Population | Land Area (Ha) | Water Area (Ha) |
|-----------------|------------|----------------|-----------------|
| Agliè           | 2,562      | 1,314.6        | 1.8             |
| Airasca         | 3,660      | 1,573.9        | 0.0             |
| Ala di Stura    | 467        | 4,633.2        | 11.0            |
| Albiano d'Ivrea | 1,637      | 1,173.1        | 0.0             |

## **2. Terrain Steepness (Slope) Classification**

The dataset categorizes land steepness into five slope classes:

-   **MOLTO BASSA (Very Low Slope)**
-   **BASSA (Low Slope)**
-   **MEDIA (Medium Slope)**
-   **ALTA (High Slope)**
-   **MOLTO ALTA (Very High Slope)**

### **Example Slope Distribution:**

| Municipality    | Very Low (%) | Low (%) | Medium (%) | High (%) | Very High (%) |
|-----------------|--------------|---------|------------|----------|---------------|
| Agliè           | 26.2         | 49.83   | 15.8       | 6.6      | 1.6           |
| Airasca         | 47.0         | 50.93   | 2.1        | 0.0      | 0.0           |
| Ala di Stura    | 0.1          | 1.29    | 6.1        | 23.1     | 69.5          |
| Albiano d'Ivrea | 48.4         | 43.05   | 7.4        | 1.1      | 0.0           |

## **3. Terrain Roughness (Rugosità)**

Roughness is measured using: - **STD SLOPE**: Standard deviation of slope. - **STD TRI**: Terrain roughness index.

Both are classified into quintiles: - **BASSA (Low Roughness)** - **MOLTO BASSA (Very Low Roughness)** - **MOLTO ALTA (Very High Roughness)**

### **Example Roughness Data:**

| Municipality    | STD SLOPE | Slope Definition | STD TRI | TRI Definition |
|-----------------|-----------|------------------|---------|----------------|
| Agliè           | 4.31      | BASSA            | 5.8     | BASSA          |
| Airasca         | 1.20      | MOLTO BASSA      | 1.5     | MOLTO BASSA    |
| Ala di Stura    | 9.47      | MOLTO ALTA       | 16.5    | MOLTO ALTA     |
| Albiano d'Ivrea | 2.17      | BASSA            | 2.8     | BASSA          |

## **4. Key Insights**

-   **Ala di Stura** has the steepest and roughest terrain, with 69.5% in "very high slope" and "very high roughness."
-   **Airasca** is the flattest municipality, with nearly all land in "very low" or "low" slope categories.
-   **Agliè** and **Albiano d'Ivrea** have moderate terrain features with mostly gentle slopes and low roughness.

This dataset provides valuable insights for urban planning, agriculture, and environmental assessment.

------------------------------------------------------------------------

## Upload Dati morfometrici

```{r,message=FALSE, warning=FALSE, echo=FALSE}
library(readxl)
library(tidyverse)
#https://www.istat.it/statistica-sperimentale/indici-per-la-descrizione-morfometrica-del-territorio-italiano/
df<- read_excel("Data/DatiMorfometrici/MorfologicData_com_2022.xlsx")
colnames(df) <- df[1, ] 
df <- df[-1, ]  
#rownames(df) <- NULL
#head(df)
head(df)
```

Upload shapefile and sorting operations

```{r, message=FALSE, warning=FALSE, echo=FALSE}

#install.packages("leaflet")
#install.packages("sf")
#install.packages("giscoR")  # To get municipality data
library(giscoR)
library(leaflet)
library(sf)
#https://public.opendatasoft.com/explore/dataset/georef-italy-comune/export/?#disjunctive.rip_code&disjunctive.rip_name&disjunctive.reg_code&disjunctive.reg_name&disjunctive.prov_code&disjunctive.prov_name&disjunctive.com_c#ode&disjunctive.com_name
# Define the path to your shapefile
shapefile_path <- "C:/Users/2692812C/OneDrive - University of Glasgow/Desktop/Projects/8 Scarface/georef-italy-comune/georef-italy-comune-millesime.shp"
# Load the shapefile
italy_comuni <- st_read(shapefile_path)
head(italy_comuni)

# Sort the shapefile dataset by municipality name
shapefile_data_sorted <- italy_comuni %>%
  arrange(com_name)  
colnames(df) <- make.unique(colnames(df))

df_sorted <- df %>%
  arrange(COMUNE)  # Sorting after fixing duplicate names
```

Plotting the data

```{r}
# Your R code here
library(dplyr)
# Ensure Municipality Names are Cleaned Before Merging
df_sorted <- df_sorted %>%
  mutate(COMUNE_clean = tolower(trimws(COMUNE)))
shapefile_data_sorted <- shapefile_data_sorted %>%
  mutate(com_name_clean = tolower(trimws(com_name)))
# Merge Based on Municipality Names
merged_data <- left_join(shapefile_data_sorted, df_sorted, 
                         by = c("com_name_clean" = "COMUNE_clean"))

# View any municipalities that did not match
unmatched <- anti_join(shapefile_data_sorted, df_sorted, by = c("com_name_clean" = "COMUNE_clean"))

head(merged_data)
str(merged_data)
merged_data$`% in ALTA`=as.numeric(merged_data$`% in ALTA`)
# Step 4: Define a Color Palette for % in ALTA
color_palette <- colorNumeric(palette = "YlOrRd", domain = merged_data$`% in ALTA`, na.color = "gray")

leaflet(data = merged_data) %>%
  addTiles() %>%
  addPolygons(fillColor = ~color_palette(`% in ALTA`),  # Color by steepness
              color = "black", weight = 0.5, opacity = 0.7,
              label = ~paste(com_name, ": ", `% in ALTA`, "%"),
              fillOpacity = 0.7) %>%
  addLegend(pal = color_palette, values = merged_data$`% in ALTA`,
            title = "% in ALTA", opacity = 1)


```

# 
